---
title: "R Notebook"
output: html_notebook
---

```{r}
library(openxlsx)
library(tidyverse)
library(synergyfinder)
library(TidyComb)
```



Jing ask me to use drugcomb curated data directly. I got a huge csv file from Jehad. Lets see how this works
```{r}
drugcomb_oneil <- read_csv("C:/Users/wenyu/Documents/work group Jing/Drugcomb/2016_ONeil/drugcomb_oneil_response.csv") 
```

search for drugpair*cell line that has highest synergy or highest antagonose 
```{r}
block_idx<- group_indices(.data = drugcomb_oneil, drug_row, drug_col, cell_line_name)
```


```{r}
res_summary <- drugcomb_oneil %>%
  bind_cols(block_idx= block_idx) %>%
  group_by(block_idx) %>%
  summarise_at(vars(synergy_zip:synergy_hsa), .funs = list(mean1=mean, median1=median, max1=max, min1=min), na.rm = TRUE) %>% ungroup()
# save(res_summary)
```


there is no such thing as 5 dose concentration in Oneil drug
```{r}
tmp <- drugcomb_oneil %>% 
  bind_cols(block_idx= block_idx) %>% 
  group_by(block_idx) %>% 
  summarize(N_conc1= length(unique(conc_r)))
```

Synergistic CI plot 7565
<!-- based on reading the Oneil paper I will first check AZD1775 (MK-1775) and MK-8776 (891494-63-6) -->
<!-- Note cell line batch replicate are all considered as replicates? -->


```{r}
# drug_row_list= sort(unique(drugcomb_oneil$drug_row))
# drug_col_list= sort(unique(drugcomb_oneil$drug_col))
data2 <- drugcomb_oneil %>% 
  bind_cols(block_idx= block_idx) %>% 
  filter(block_idx==12737) %>%
  # filter(drug_row=="5-Fluorouracil") %>%
  # filter(drug_col == "Veliparib") %>%
  # filter(cell_line_name== "A2058") %>%
  mutate(ConcUnit= "uM") %>% 
  mutate(block_id =1 ) 
# write_csv(data2, "Oneil_synergistic_testdata.csv")


data2_matrix_format <-  data2 %>% 
  select(conc_r, conc_c, inhibition) %>% 
  group_by(conc_r,conc_c) %>% 
  summarise(inhibition = mean(inhibition)) %>% 
  pivot_wider(names_from = conc_c, values_from = inhibition)

write_csv(data2_matrix_format, "Oneil_synergistic_testdata_matrix_format.csv")


res_response <- ReshapeData(data2, data_type = "inhibition")
res_HSA <- CalculateSynergy(res_response, method = c("HSA"))
res_Bliss <- CalculateSynergy(res_response, method = c("Bliss"))
res_Loewe <- CalculateSynergy(res_response, method = c("Loewe"))
res_ZIP <- CalculateSynergy(res_response, method = c("ZIP"))
res_list <- list(response=res_response, HSA_synergy= res_HSA, Bliss_synergy=res_Bliss, Loewe_synergy=res_Loewe, ZIP_synergy= res_ZIP)

Plot2DrugHeatmap(res_response,plot_value = "response", plot_block = 1, statistic = "ci",summary_statistic = "mean") 

for (i in 1:5){
  res <- res_list[[i]]
  p <- Plot2DrugHeatmap(res,plot_value = names(res_list[i]), plot_block = 1, statistic = "ci",summary_statistic = "mean") +
  theme(axis.title.x=element_text(size = 12))+
  theme(axis.text.x = element_text(size = 12,color = "black"))+
  theme(axis.title.y=element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12,color = "black"))
  # print(p)
  ggsave(p, filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/ciplot","/synergistic.", i,".pdf"), dpi=500,width = 18,height =8,units = "cm" )
}
```


```{r}
  res <- CalculateSynergy(res_response, method = c("HSA","Bliss","Loewe", "ZIP"))
  p <- PlotBarometer(res,plot_concs = c(0.200,2.750),needle_text_size = 2.8,annotation_label_size = 5,needle_text_offset = -2,graduation_label_size = 2.8)
  print(p)
  ggsave(p, filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/baro_synergistic.pdf"), dpi=500,width = 8,height =8,units = "cm" )


```



```{r}
# Antagonistic drug pair
data2 <- drugcomb_oneil %>% 
  bind_cols(block_idx= block_idx) %>% 
  filter(block_idx== 17843) %>% 
  mutate(ConcUnit= "uM") %>% 
  mutate(block_id =1 )
data2$drug_row <- "Paclitaxel"

write_csv(data2, "Oneil_antagonistic_testdata.csv")

data2_matrix_format <-  data2 %>% 
  select(conc_r, conc_c, inhibition) %>% 
  group_by(conc_r,conc_c) %>% 
  summarise(inhibition = mean(inhibition)) %>% 
  pivot_wider(names_from = conc_c, values_from = inhibition)

write_csv(data2_matrix_format, "Oneil_antagonistic_testdata_matrix_format.csv")


library(synergyfinder)
res_response <- ReshapeData(data2, data_type = "inhibition")
res_hsa <- CalculateSynergy(res_response, method = c("HSA"))
res_Bliss <- CalculateSynergy(res_response, method = c("Bliss"))
res_loewe <- CalculateSynergy(res_response, method = c("Loewe"))
res_zip <- CalculateSynergy(res_response, method = c("ZIP"))
res_list <- list(response=res_response, HSA_synergy= res_hsa, Bliss_synergy=res_Bliss, Loewe_synergy=res_loewe, ZIP_synergy= res_zip)

for (i in 1:5){
  res <- res_list[[i]]
  p <- Plot2DrugHeatmap(res,plot_value = names(res_list[i]), plot_block = 1, statistic = "ci",summary_statistic = "mean") +
  theme(axis.title.x=element_text(size = 12))+
  theme(axis.text.x = element_text(size = 12,color = "black"))+
  theme(axis.title.y=element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12,color = "black"))
  ggsave(p, filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/ciplot","/antagonistic.", i,".pdf"), dpi=500,width = 18,height =8,units = "cm" )
}
```



```{r}
res <- CalculateSynergy(res_response, method = c("HSA","Bliss","Loewe", "ZIP"))
tmp <- res[["synergy_scores"]] %>% select(conc1,conc2, contains("ref"))
p <- PlotBarometer(res,plot_concs = c(0.001600, 2.000),needle_text_size = 2.8,annotation_label_size = 5,needle_text_offset = -2,graduation_label_size = 2.8)
p
ggsave(p, filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/baro_antagonistic.pdf"), dpi=500,width = 8,height = 8.2,units = "cm" )
```


SS plot
```{r}
# for this sensitivity-synergy plot you need multiple datapoints rather than only one drug pair
# note there is a type in the function help page PlotSensitiveSynergy (Loewe not leowe)
# there is also typo in the CalculateSensitivity

# it will be superslow to use replicates. Use median value instead.
# alternatively I can construct a synergy object by my own using existing synergy scores. 
# The idea is to test PlotSensitiveSynergy function

#get clear what input data is wanted for this function as data is now quite large.
block_idx<- group_indices(.data = drugcomb_oneil, drug_row, drug_col)
data2 <- drugcomb_oneil %>% 
  bind_cols(block_idx= block_idx) %>%
  select(-block_id) %>% 
  rename(block_id=block_idx) %>% 
  mutate(ConcUnit= "uM") %>% 
  # filter(block_id<3) %>%
  filter(cell_line_name== "A2058") %>% 
  select(-synergy_zip, -synergy_bliss, -synergy_loewe,-synergy_hsa) %>% 
  group_by_at(-5) %>% 
  summarise(inhibition= median(inhibition)) %>% 
  ungroup()
  
res_oneil <-  ReshapeData(data2, data_type = "inhibition") 
res_oneil <-  CalculateSensitivity(res_oneil) 
save(res_oneil,file = "res_oneil.RData")
load("res_oneil.RData")
oneil_hsa <-  CalculateSynergy(res_oneil, method = c("HSA"))
oneil_bliss <-  CalculateSynergy(res_oneil, method = c("Bliss"))
oneil_loewe <-  CalculateSynergy(res_oneil, method = c("Loewe"))
oneil_ZIP<-  CalculateSynergy(res_oneil, method = c("ZIP"))
# #there should be a parameter allowing user to choose whether they want to add geom_text.
save(list = c("res_oneil", "oneil_bliss", "oneil_hsa", "oneil_loewe", "oneil_ZIP"),file = "res_oneil.RData")
```


```{r}
PlotSensitiveSynergy(oneil_hsa,plot_synergy = "HSA",point_size = 0.5)+
  ylab("Synergy Score")+
    theme(axis.title.x=element_text(size = 10))+
  theme(axis.text.x = element_text(size = 10,color = "black"))+
  theme(axis.title.y=element_text(size = 10))+
  theme(axis.text.y = element_text(size = 10,color = "black"))+
  ggtitle("HSA")+
  theme(plot.margin=unit(c(t=1,b=1,l=1,r=1),"cm"))+
  ggsave(filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/hsa.pdf"), dpi=500,width = 8,height =8,units = "cm" )
  
  
PlotSensitiveSynergy(oneil_loewe,plot_synergy = "Loewe",point_size = 0.5)+
    ylab("Synergy Score")+
    theme(axis.title.x=element_text(size = 10))+
  theme(axis.text.x = element_text(size = 10,color = "black"))+
  theme(axis.title.y=element_text(size = 10))+
  theme(axis.text.y = element_text(size = 10,color = "black"))+
  ggtitle("Loewe")+
  theme(plot.margin=unit(c(t=1,b=1,l=1,r=1),"cm"))+
  ggsave(filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/loewe.pdf"), dpi=500,width = 8,height =8,units = "cm" )
  
PlotSensitiveSynergy(oneil_bliss,plot_synergy = "Bliss",point_size = 0.5)+
      ylab("Synergy Score")+
    theme(axis.title.x=element_text(size = 10))+
  theme(axis.text.x = element_text(size = 10,color = "black"))+
  theme(axis.title.y=element_text(size = 10))+
  theme(axis.text.y = element_text(size = 10,color = "black"))+
  ggtitle("Bliss")+
  theme(plot.margin=unit(c(t=1,b=1,l=1,r=1),"cm"))+
  ggsave(filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/bliss.pdf"), dpi=500,width = 8,height =8,units = "cm" )

PlotSensitiveSynergy(oneil_ZIP,plot_synergy = "ZIP",point_size = 0.5)+
        ylab("Synergy Score")+
    theme(axis.title.x=element_text(size = 10))+
  theme(axis.text.x = element_text(size = 10,color = "black"))+
  theme(axis.title.y=element_text(size = 10))+
  theme(axis.text.y = element_text(size = 10,color = "black"))+
  ggtitle("ZIP")+
  theme(plot.margin=unit(c(t=1,b=1,l=1,r=1),"cm"))+
  ggsave(filename = paste0("C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig5/zip.pdf"), dpi=500,width = 8,height =8,units = "cm" )


```


Figure Z. A) An example table of a drug combination. B) Main targets/classifications/clinical stages of the O'neill drugs
```{r}
library(TidyComb)
block_idx<- group_indices(.data = drugcomb_oneil, drug_row, drug_col)
data2 <- drugcomb_oneil %>% 
  bind_cols(block_idx= block_idx) %>%
  select(-block_id) %>% 
  rename(block_id=block_idx) %>% 
  mutate(ConcUnit= "uM") %>% 
  filter(block_id<2) %>%
  filter(cell_line_name== "A2058") %>% 
  select(-synergy_zip, -synergy_bliss, -synergy_loewe,-synergy_hsa) %>% 
  group_by_at(-5) %>% 
  summarise(inhibition= median(inhibition)) %>% 
  ungroup()
# now you have a 5by5 viability datatable for   one drug pair in A2058 cell line

```

```{r}
drug_row_list= sort(unique(drugcomb_oneil$drug_row))
drug_col_list= sort(unique(drugcomb_oneil$drug_col))
tmp_cell <- AnnotateCell(cell_names = unique(drugcomb_oneil$cell_line_name),
                         file = "C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/data/cell_anno/tmp.xml")
tmp_cell$tissue[is.na(tmp_cell$tissue)]= "Unknown"  
tmp <- AnnotateDrug(drug_row_list)
tmp1 <- tmp[["drug_annotation"]]

tmp2 <- tmp[["target"]] 
tmp2 <- tmp2 %>% 
  separate_rows(target_protein_class,sep = ",") %>%
  mutate(target_protein_class= str_trim(.data$target_protein_class)) %>% 
  select(-inhibitor_type)
tmp2$target_protein_class[tmp2$target_protein_class==""]= "Unknown"
tmp2$target_protein_class[is.na(tmp2$target_protein_class)]= "Unknown"  



# DownloadCellosaurus(file = "C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/data/cell_anno/tmp.xml")
# View(AnnotateDrug("5-Fluorouracil")[["drug_annotation"]])

tmp_figure6a <- drugcomb_oneil %>% 
    filter(cell_line_name== "A2058" ) %>% 
    filter(drug_row== "5-Fluorouracil")%>%
    filter(drug_col=="Vorinostat") %>% 
    mutate(block_id= 1) %>% 
    mutate(conc_unit= "um") %>% 
    mutate(inhibition= round(.data$inhibition, digits = 2)) %>% 
    select(cell_line_name, drug_row,conc_r,drug_col,conc_c,conc_unit, inhibition, block_id)
  

write_csv(tmp_figure6a, "C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/res_fig/fig6/data_app.csv")

View(drugcomb_oneil %>% 
       filter(cell_line_name== "A2058" ) %>% 
       filter(drug_row== "5-Fluorouracil")%>%
       filter(drug_col=="Vorinostat") %>%
       select_at(c(2:6,11)) %>% 
       mutate(conc_unit= "um") %>% 
       mutate(inhibition= round(.data$inhibition, digits = 2)) %>% 
       select(cell_line_name, drug_row,conc_r,drug_col,conc_c,conc_unit, inhibition))
View(tmp1 %>% filter(input_name == "5-Fluorouracil") %>% select(-smiles,-inchikey))
View(tmp1 %>% filter(input_name == "Vorinostat") %>% select(-smiles, -inchikey))
View(tmp2 %>% filter(chembl_id== "CHEMBL185"))
View(tmp2 %>% filter(chembl_id== "CHEMBL98"))
tmp5 <- AnnotateCell(data2$cell_line_name[1],file = "C:/Users/wenyu/Documents/work group Jing/Drugcomb/synergyfinder/synergyfinderplus_showcase/data/cell_anno/tmp.xml")


```



```{r}
library(scales)
clinical_phase_freq <- tmp1 %>% 
  group_by(clinical_phase) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n),
         text = paste0(round(freq, 4) * 100, " %")
         ) %>% 
  arrange(freq)

tmp1$clinical_phase <- factor(
  tmp1$clinical_phase,
  levels = c("", 4, 3, 2, 1),
  labels = c("Unknown", "Phase IV", "Phase III", "Phase II", "Phase I"))
clinical_phase_freq$clinical_phase <- factor(
  clinical_phase_freq$clinical_phase,
  levels = c("", 4, 3, 2, 1),
  labels = c("Unknown", "Phase IV", "Phase III", "Phase II", "Phase I"))

clinical_phase_plot <- ggplot(data = tmp1, aes(y = clinical_phase)) +
  geom_bar(fill = "#5D82BD") +
  geom_text(data = clinical_phase_freq, aes(x = n, y = clinical_phase, label = text),
            nudge_x = 0.3, size = 9/.pt) +
  theme_classic() +
  theme(axis.title.x=element_text(size = 12))+
  theme(axis.text.x = element_text(size = 12,color = "black"))+
  theme(axis.title.y=element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12,color = "black"))+
  theme(axis.title.y = element_blank()) +
  # xlim(c(0,4500)) +
  scale_x_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)),
                     limits = c(1, 10^2)#,
                     # expand = c(10, 10)
                     )  +
  # scale_x_continuous(limits = c(0, 4500), expand = c(0, 1)) +
  ggtitle("Clinical Trial Phase")# +
  # theme(
  #   plot.title = element_text(hjust = 0.5),
  #   axis.title.y = element_blank(),
  #   axis.text.y = element_text(margin = margin(r = 0)),
  #   panel.background = element_blank(),
  #   panel.grid.major.y = element_blank(),
  #   panel.grid.minor.y = element_blank(),
  #   panel.grid.major.x = element_line(color = "grey90"),
  #   panel.grid.minor.x = element_line(color = "grey90"),
  #   axis.ticks = element_blank()
  #   )
clinical_phase_plot
ggsave(clinical_phase_plot, filename = "figure6_b1.pdf" ,dpi=500,
       width = 8,
       height = 5.5,units = "cm" )
```

Target protein category
```{r}
library(scales)
target_protein_category_freq <- tmp2 %>% 
  group_by(target_protein_class) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n),
         text = paste0(round(freq, 4) * 100, " %")
         ) %>% 
  arrange(desc(freq))

tmp2$target_protein_class <- factor(
  tmp2$target_protein_class,
  levels = target_protein_category_freq$target_protein_class,
  labels = target_protein_category_freq$target_protein_class
  )
target_protein_category_freq$target_protein_class <- factor(
  target_protein_category_freq$target_protein_class,
  levels = target_protein_category_freq$target_protein_class,
  labels = target_protein_category_freq$target_protein_class
  )

target_protein_category_plot <- ggplot(data = tmp2, aes(y = target_protein_class)) +
  geom_bar(fill = "#5D82BD") +
  geom_text(data = target_protein_category_freq, aes(x = n, y = target_protein_class, label = text),
            nudge_x = 0.32, size = 9/.pt) +
  theme_classic() +
  theme(axis.title.x=element_text(size = 12))+
  theme(axis.text.x = element_text(size = 12,color = "black"))+
  theme(axis.title.y=element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12,color = "black"))+
  theme(axis.title.y = element_blank()) +
  # xlim(c(0,4500)) +
  scale_x_continuous(trans = "log10",
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)),
                     limits = c(1, 10^2.5)#,
                     # expand = c(10, 10)
                     )  +
  # scale_x_continuous(limits = c(0, 4500), expand = c(0, 1)) +
  ggtitle("Target Protein Class")# +
  # theme(
  #   plot.title = element_text(hjust = 0.5),
  #   axis.title.y = element_blank(),
  #   axis.text.y = element_text(margin = margin(r = 0)),
  #   panel.background = element_blank(),
  #   panel.grid.major.y = element_blank(),
  #   panel.grid.minor.y = element_blank(),
  #   panel.grid.major.x = element_line(color = "grey90"),
  #   panel.grid.minor.x = element_line(color = "grey90"),
  #   axis.ticks = element_blank()
  #   )
target_protein_category_plot
ggsave(target_protein_category_plot,
       filename = "figure6_b2.pdf" ,dpi=500,
       width = 10,
       height =8,units = "cm" )
```
cell line tissue type
```{r}
library(scales)
tissue_freq <- tmp_cell %>% 
  group_by(tissue) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n),
         text = paste0(round(freq, 4) * 100, " %")
         ) %>% 
  arrange(desc(freq))

tmp_cell$tissue <- factor(
  tmp_cell$tissue,
  levels = tissue_freq$tissue,
  labels = c( "Unknown","Ovary","Lung", "Large_intestine","Skin","Breast",  "Prostate"  )
  )
tissue_freq$tissue <- factor(
  tissue_freq$tissue,
  levels = tissue_freq$tissue,
  labels =c( "Unknown","Ovary","Lung", "Large_intestine","Skin","Breast",  "Prostate"  )
  )

tissue_plot <- ggplot(data = tmp_cell, aes(y = tissue)) +
  geom_bar(fill = "#5D82BD") +
  geom_text(data = tissue_freq, aes(x = n, y = tissue, label = text),
            nudge_x = 1.8, size = 9/.pt) +
  theme_classic() +
  theme(axis.title.x=element_text(size = 12))+
  theme(axis.text.x = element_text(size = 12,color = "black"))+
  theme(axis.title.y=element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12,color = "black"))+
  theme(axis.title.y = element_blank()) +
  xlim(c(0,13)) +
  # scale_x_continuous(trans = "log10",
  #                    breaks = trans_breaks("log10", function(x) 10^x),
  #                    labels = trans_format("log10", math_format(10^.x)),
  #                    limits = c(1, 10^1.3)#,
  #                    # expand = c(10, 10)
  #                    )  +
  # scale_x_continuous(limits = c(0, 4500), expand = c(0, 1)) +
  ggtitle("Cell Line Tissue Origin")# +
  # theme(
  #   plot.title = element_text(hjust = 0.5),
  #   axis.title.y = element_blank(),
  #   axis.text.y = element_text(margin = margin(r = 0)),
  #   panel.background = element_blank(),
  #   panel.grid.major.y = element_blank(),
  #   panel.grid.minor.y = element_blank(),
  #   panel.grid.major.x = element_line(color = "grey90"),
  #   panel.grid.minor.x = element_line(color = "grey90"),
  #   axis.ticks = element_blank()
  #   )
tissue_plot

ggsave(tissue_plot,
       filename = "figure6_b3.pdf" ,dpi=500,
       width = 9,
       height =6.5,units = "cm" )

 tissue_freq$tissue
```


